---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo=FALSE}
knitr::opts_knit$set(root.dir = "./inst/extdata/") 
```

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

Current Travis Build Status: [![Travis-CI Build Status](https://travis-ci.org/RussellPolitzky/fars.svg?branch=master)](https://travis-ci.org/RussellPolitzky/fars)

# FARS Analysis Functions

FARS is the Fatal Accident Reporting System.  Fatal Accident Reports provide details, such as the time and location of each accident.  The `fars` package provides a set of simple functions to load and analyze this data.

## `make_file_name`

FARS, CSV data files have names with the following form:

    accident_xxxx.csv.bz2

Where `xxxx` is the year expressed as a four digit number.  For example, the FARS file for 2013 is:

    accident_2013.csv.bz2

Givent the year, `make_file_name()` generates this file name:

```{r}
library(fars)

fars_file <- make_filename(2013)
fars_file
```

## `fars_read()`

Given the year of interest, `fars_read()` will read the associated FARS, CSV data file into a `data.frame`.




```{r }
fars_file <- make_filename(2013)
fars_data <- fars_read(fars_file)
fars_data
```

## `fars_read_years()`

Most users will want to load several FARS data files, by year, without having to generate the associated file names.  `fars_read_years()` makes this possible.  The function takes a vector of years and a list of `data.frame`s; one for each given year.

```{r eval=FALSE}
all_data <- fars_read_years(c(2013, 2014, 2015))
all_data
```

The resulting list of `data.frame`s may, be subsequently combined, and analyzed as a whole.

## `fars_summarize_years()`

This function generates a summary of fatalities per month, for the given vector of years. `fars_summarize_years()` takes a single argument; years; a vector/list of years that the summary will cover.

```{r}
fars_summarize_years(c(2013, 2014, 2015))
```

## `fars_map_state()`

Given a state and a year, `fars_map_state` plots the location of each fatal accident which ocurred in that year.  The density of plotted dots is an indication of the number of fatal accidents in a given area.

```{r}
library(maps)
data(stateMapEnv)
fars_map_state(51, 2013)
```






